\documentclass[10.5pt,a4paper]{article}

% ---------- Packages (all standard, ATS-safe: no icons, no columns, no tables) ----------
\usepackage[margin=0.55in,top=0.5in,bottom=0.5in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{mathptmx}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[skip=3pt]{parskip}

% ---------- Colors ----------
\definecolor{headcolor}{RGB}{20,20,20}
\definecolor{linkcolor}{RGB}{0,0,0}

% ---------- Hyperlink setup (plain, no colored boxes -> ATS/printer safe) ----------
\hypersetup{
    colorlinks=true,
    linkcolor=linkcolor,
    urlcolor=linkcolor,
}

% ---------- Section formatting ----------
\titleformat{\section}{\normalsize\bfseries\color{headcolor}}{}{0em}{}[\vspace{1pt}\titlerule\vspace{1pt}]
\titlespacing{\section}{0pt}{6pt}{2pt}

% ---------- List formatting ----------
\setlist[itemize]{leftmargin=14pt, topsep=1pt, itemsep=1pt, parsep=0pt}

\setlength{\parindent}{0pt}
\pagestyle{empty}

% ---------- Custom commands ----------
\newcommand{\resumeItem}[1]{\item #1}

\newcommand{\resumeEntry}[3]{
  \textbf{#1} \hfill \textit{#2} \\[-1pt]
  {\small #3}
}

\begin{document}

% ============================ HEADER ============================
\begin{center}
    {\Huge \bfseries Shubham Kumar}\\[6pt]
    \small
    +91-7667956792 ~ $\vert$ ~
    \href{mailto:rishab.ss64@gmail.com}{rishab.ss64@gmail.com} ~ $\vert$ ~
    \href{https://github.com/shubham7667}{github.com/shubham7667} ~ $\vert$ ~
    Ranchi, Jharkhand
\end{center}

\vspace{2pt}

% ============================ SUMMARY ============================
\section*{Professional Summary}
M.Tech candidate in Artificial Intelligence & Machine Learning at Birla Institute of Technology, Mesra with hands-on experience in NLP, machine learning, and data science. Proficient in building end-to-end ML pipelines using Python, Scikit-learn, NumPy, and Pandas, with practical experience in deep learning using PyTorch and LLM application development using LangChain. Strong foundation in data structures, algorithms, and core computer science concepts, with 150+ solved coding problems. Passionate about building intelligent, data-driven systems and seeking AI/ML or software internships to apply and expand technical expertise.


% ============================ EDUCATION ============================
\section*{Education}

\textbf{M.Tech -- Artificial Intelligence \& Machine Learning} \hfill CGPA: 8.1 (Ongoing) \\
Birla Institute of Technology (BIT) Mesra \hfill 2025 -- 2027 \\
GATE 2025 Qualified --- CSIT Stream

\vspace{2pt}

\textbf{B.Tech -- Computer Science \& Engineering} \hfill CGPA: 7.79 \\
Sarala Birla University, Ranchi \hfill 2021 -- 2024

% ============================ SKILLS ============================
\section*{Technical Skills}
\begin{itemize}[leftmargin=14pt, topsep=2pt, itemsep=2pt]
    \item \textbf{Languages:} Python, C
    \item \textbf{ML / DL Libraries:} Scikit-learn, NumPy, Pandas, PyTorch (basics), NLTK
    \item \textbf{NLP \& LLM:} Transformers (BERT), LangChain, Retrieval-Augmented Generation (RAG), FAISS, Word2Vec, TF-IDF
    \item \textbf{Core CS:} Data Structures \& Algorithms, DBMS, Operating Systems, OOP
    \item \textbf{Tools \& Platforms:} Git, GitHub, Streamlit, Jupyter Notebook, VS Code, Google Colab
    \item \textbf{Domains:} Machine Learning, NLP, Evolutionary Computing, Data Analysis, Clustering, Optimization
\end{itemize}

% ============================ PROJECTS ============================
\section*{Projects \& Academic Work}

\textbf{Word Sense Disambiguation using Lesk Algorithm} \hfill \textit{Python, NLTK, WordNet}
\begin{itemize}
    \item Designed an NLP pipeline to resolve lexical ambiguity in text by leveraging the simplified Lesk algorithm with WordNet lexical databases.
    \item Implemented text preprocessing (tokenization, stop-word removal, lemmatization) to improve contextual overlap accuracy by ~18\%.
    \item Evaluated performance on SemEval benchmark datasets; outcome demonstrates applicability in chatbot and IR systems.
\end{itemize}

\vspace{2pt}
\textbf{LULC Analysis using K-Means Clustering (Remote Sensing)} \hfill \textit{Python, Scikit-learn, Rasterio, Sentinel-2 Imagery}
\begin{itemize}
    \item Built an unsupervised land-use/land-cover classification system processing multispectral Sentinel-2 satellite imagery using K-Means clustering.
    \item Applied spectral band engineering and image normalization to distinguish vegetation, water bodies, urban zones, and barren land.
    \item Visualized temporal LULC changes across multiple years to support environmental monitoring use-cases.
\end{itemize}

\vspace{2pt}
\textbf{Sentiment Analysis on Product Reviews} \hfill \textit{Python, NLTK, Scikit-learn, TF-IDF}
\begin{itemize}
    \item Developed a binary sentiment classifier (positive/negative) trained on e-commerce review datasets using TF-IDF features and Logistic Regression.
    \item Achieved 88\%+ accuracy with cross-validated evaluation; explored Naive Bayes and SVM baselines for comparison.
\end{itemize}

\vspace{2pt}
\textbf{DocChat -- Retrieval-Augmented Generation (RAG) System for PDF Q\&A} \hfill \textit{Python, LangChain, FAISS, HuggingFace, Streamlit, Gemini}
\begin{itemize}
    \item Built an end-to-end modular RAG pipeline (Document Loader $\rightarrow$ Chunking $\rightarrow$ Embedding $\rightarrow$ Vector Store $\rightarrow$ Retriever $\rightarrow$ LLM) enabling natural-language question answering over PDF documents.
    \item Implemented configurable recursive text chunking and semantic embeddings (sentence-transformers/all-MiniLM-L6-v2) indexed with FAISS for top-k similarity retrieval.
    \item Integrated Google Gemini 2.5 Flash via LangChain with a context-grounded prompt template to minimize hallucination, and built an interactive Streamlit interface with adjustable chunking parameters, pipeline-status tracking, and retrieved-context visualization.
\end{itemize}

% ============================ ACHIEVEMENTS ============================
\section*{Achievements \& Co-Curriculars}
\begin{itemize}
    \item Cleared GATE 2025 in AI/ML stream, demonstrating mastery of core CS and ML theory.
    \item Actively solving Striver A2Z DSA Sheet --- 150+ problems across arrays, trees, graphs, and DP.
    \item Completed coursework in Evolutionary Computing covering GA, NSGA-II, MCTS, and multi-objective optimization.
    \item Contributed to a group research project implementing Niching GA, Multi-Objective GA, and Competitive Coevolution for job-scheduling optimization with unified visualizations.
    
\end{itemize}

\end{document}
